### 2025-04-14 paper/paper.qmd
Added changes for thematic analysis

---
title: "STA107 Post-Course Survey Analysis"
subtitle: "Evaluating Student Reflections on R and Statistical Learning"
author: 
  - Sinan Ma
  - Jaiditya Dev
date: "April 13, 2025"
format: 
  pdf:
    toc: true
    number-sections: true
    code-fold: true
    code-summary: "Show Code"
    fig-align: center
    fig-cap-location: top
    keep-tex: true
execute:
  echo: false  
  warning: false
  message: false
---

```{r}
#| include: false
#| warning: false
#| message: false

#### workspace setup ####
library(tidyverse)
library(readr)
library(janitor)
library(here)
library(ggplot2)
```

```{r setup, include=FALSE}

#### import survey data ####
data <- read_csv("../data/clean_data.csv")
qualitative_data <- read_csv("../data/qualitative_analysis_data.csv")
quantitative_data <- read_csv("../data/quantitative_analysis_data.csv")
```

# Introduction

This report presents an analysis of the anonymous post-course survey completed by students enrolled in **STA107: Introduction to Statistics** at the University of Toronto Mississauga. The survey was voluntary, open for multiple submissions, and contributed **2% to the course grade**. Its goal was to evaluate students’ experiences with R-based activities, the integration of normal distribution concepts, and overall satisfaction with the course components. This feedback will help refine future course offerings and improve the learning experience.

# Survey Questions

This section outlines the structure of the survey and highlights key question types:

- **Quantitative items**: Likert-scale responses on course clarity, usefulness of R exercises, and understanding of statistical concepts.
- **Qualitative items**: Open-ended prompts for students to share what they found most/least helpful and suggestions for improvement.

The post-course survey comprised 17 open- and close-ended questions. Among these, 8 were qualitative in nature (`Q3`, `Q6`, `Q8`, `Q10`, `Q11`, `Q12`, `Q13`, `Q14`, `Q15`, `Q16`, `Q17`), focusing on free-text responses reflecting students' thoughts on the content, instruction, and overall experience.

# Data Analysis

## Methodology

We employed a **mixed-methods approach**:

- **Quantitative analysis**: Descriptive statistics and visualizations for Likert-scale questions.
- **Qualitative analysis**: Thematic coding of open-ended responses to identify common sentiments and suggestions.

## Quantitative Analysis
```{r}
# Clean column names
colnames(quantitative_data) <- colnames(quantitative_data) %>%
  tolower() %>%
  str_replace_all(" ", "_") %>%
  str_replace_all("^x", "")

# Select and rename Likert-scale columns
quantitative_data <- quantitative_data[, 1:6]
colnames(quantitative_data) <- c(
  "Stat_Communication_Before_Course",
  "Stat_Communication_After_Course",
  "R_Skills_Before_Course",
  "R_Skills_After_Course",
  "Understanding_After_Assignments",
  "Understanding_After_Tutorials"
)
```

### Communication Ability Before vs. After
```{r}
comm_data <- quantitative_data %>%
  select(Stat_Communication_Before_Course, Stat_Communication_After_Course) %>%
  pivot_longer(cols = everything(), names_to = "Question", values_to = "Rating") %>%
  drop_na()

ggplot(comm_data, aes(x = factor(Rating), fill = Question)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(title = "Ability to Communicate Statistics: Before vs. After",
       x = "Rating (1 = Lowest, 5 = Highest)", y = "Count") +
  scale_fill_manual(values = c("Stat_Communication_Before_Course" = "#66c2a5", "Stat_Communication_After_Course" = "#fc8d62"))
```

The histogram shows a clear improvement in students' perceived communication ability. More students rated themselves at level 4 or 5 after the course compared to before, indicating that the course helped them gain confidence in articulating statistical ideas.

### R Ability Before vs. After
```{r}
r_data <- quantitative_data %>%
  select(R_Skills_Before_Course, R_Skills_After_Course) %>%
  pivot_longer(cols = everything(), names_to = "Question", values_to = "Rating") %>%
  drop_na()

ggplot(r_data, aes(x = factor(Rating), fill = Question)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(title = "Ability to Use R: Before vs. After",
       x = "Rating (1 = Lowest, 5 = Highest)", y = "Count") +
  scale_fill_manual(values = c("R_Skills_Before_Course" = "#8da0cb", "R_Skills_After_Course" = "#e78ac3"))
```

Students entered the course with relatively low self-reported R skills (mostly 1s and 2s), but after completing the course, a majority rated themselves at 3 or higher. This shift demonstrates the course’s success in introducing R programming effectively.

### Understanding After Assignments vs. Tutorials
```{r}
concept_data <- quantitative_data %>%
  select(Understanding_After_Assignments, Understanding_After_Tutorials) %>%
  pivot_longer(cols = everything(), names_to = "Question", values_to = "Rating") %>%
  drop_na()

ggplot(concept_data, aes(x = factor(Rating), fill = Question)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(title = "Understanding of Statistical Concepts: Assignment vs. Tutorial",
       x = "Rating (1 = Lowest, 5 = Highest)", y = "Count") +
  scale_fill_manual(values = c("Understanding_After_Assignments" = "#a6d854", "Understanding_After_Tutorials" = "#ffd92f"))
```

Students found both tutorials and assignments useful for understanding statistical concepts, with a slight preference for tutorials. The higher number of level 4 and 5 responses for tutorials suggests that real-time, interactive support may have been especially beneficial.

## Qualitative Analysis
```{r}
# Load required libraries
library(tidyverse)
library(stringr)
library(tidytext)
library(syuzhet)
library(ggplot2)
library(tidyr)
library(wordcloud)

# Read in the survey data
data <- qualitative_data

# Rename columns
colnames(data)[1:8] <- c("Q3", "Q6", "Q8", "Q10", "Q13", "Q15", "Q16", "Q17")

# Clean the text
data_clean <- data %>%
  mutate(across(everything(), ~ str_trim(tolower(as.character(.)))))

# Initialize an empty data frame to collect sentiment results
sentiment_results <- data.frame(
  Question = character(),
  Positive = numeric(),
  Negative = numeric(),
  Neutral = numeric(),
  stringsAsFactors = FALSE
)

# Function to calculate sentiment
analyze_sentiment <- function(question_col, question_name) {
  sentiments <- get_nrc_sentiment(data_clean[[question_col]])

  pos <- sum(sentiments$positive)
  neg <- sum(sentiments$negative)
  neutral <- nrow(sentiments) - 
    sum(rowSums(sentiments[, c("positive", "negative")]) > 0)

  sentiment_results <<- rbind(sentiment_results, data.frame(
    Question = question_name,
    Positive = pos,
    Negative = neg,
    Neutral = neutral
  ))
}

# Apply the function to all relevant questions
question_names <- c("Q3", "Q6", "Q8", "Q10", "Q13", "Q15", "Q16", "Q17")

for (qn in question_names) {
  analyze_sentiment(qn, qn)
}
```
# Results

## Quantitative Results

The survey included six Likert-scale questions aimed at capturing students’ self-perceived development in statistical understanding and communication, both before and after the course. The questions also addressed students’ ability to use R software and their comprehension of statistical concepts through tutorials and weekly assignments.

Key findings include:

- **Improved Communication Skills**: Students reported a notable improvement in their ability to communicate statistics, with the average post-course rating (M = 3.92) higher than the pre-course rating (M = 3.09).
- **Increased R Proficiency**: The mean rating for R software usage rose from 2.43 before the course to 3.49 after completing it, indicating significant growth in computational skills.
- **Conceptual Understanding**: Understanding of statistical concepts improved through both tutorials (M = 4.09) and weekly assignments (M = 3.84), with tutorials being rated slightly more helpful on average.

The graphs above illustrate the distribution of responses across the three themes: communication, R proficiency, and statistical understanding.

## Qualitative Results

To gain deeper insights into student opinions, the qualitative analysis comprised two parts: thematic analysis and sentiment analysis.

### Thematic Analysis

Following Braun and Clarke’s (2006) framework, a sample of approximately 80 student responses was examined through six phases:

1. **Familiarization**: Students’ responses across 8 open-ended questions were read and annotated to gain an overall understanding.

2. **Generating Initial Codes**: Semantic-level codes were created to capture features related to course experience, R skill development, and content relevance.

3. **Searching for Themes**: Related codes were clustered into preliminary themes reflecting student attitudes and feedback.

4. **Reviewing Themes**: Themes were refined to ensure clarity and distinctiveness across the dataset.

5. **Defining and Naming Themes**: Clear definitions were established for each theme based on recurring patterns.

6. **Producing the Report**: Final themes were reported below with supporting quotes.

### Identified Themes

#### Relevance to Real Life and Careers
Many students expressed appreciation for statistics content when it connected to real-world examples or their field of study.

> "This course gave me a solid foundation in statistical concepts and taught me how to apply them using R. Overall, I gained practical skills and became more confident in working with data."

> "The course showed me how statistics applies to real-world problems, such as data analysis in my ENV course as it has alot of data."

#### R as a Learning Tool
Students had mixed responses about learning R. While some found it helpful, others found it initially intimidating.

> "After the course, I had a higher opinion of my ability to use R because I learned how to create visualizations and perform statistical analyses such as normal distribution applications and regression. Now I can use it to organize data, generate graphs, and interpret the output. The tutorial session and assignments were particularly helpful in helping me gradually build my skills."

> "I have never worked with R, but after trying to work with it I realized that it is also a powerful language."

#### Clarity and Structure of Assignments
Clearer instructions and expectations were a recurring theme.

> "The weekly assignment was a R assignment based on the module. It was sometimes hard to understand the purpose of the task."

> "After completing the weekly assignments, I have a general understanding of the statistical concepts in each module, but sometimes I still have some difficulties understanding some difficult concepts."

#### Appreciation for Tutorials and TA Support
Tutorials and TA help sessions were viewed positively, especially for R and concept reinforcement.

> "For each question on the tutorial worksheet, my TA explained the relevant concepts before attempting the problem, which helped me understand the concept better. In addition, all tutorial exercises were related to the weekly assignment, so it gave me reassurance that I was doing everything correctly before submitting my assignments."

> "The weekly tutorial was extremely useful for me. The worksheets were helpful and my TA, Dylan, did an excellent job at explaining the concepts involved and answering our questions when we were confused."

> "my ta was great at explaining concepts and let us ask many questions which would all get responses. very good by jaiditya."

#### Feedback and Reflection
Students wanted more timely or individualized feedback.

> "I always read the feedback give by the TA's since they allowed me to understand how to communicate my answers better. The statistical examples did help me recognize the usefulness and value of learning about statistics because it made me realize that statistics is everywhere."

> "Most of the feedback I recieved was in the form of checks or X. I feel like I could have gotten a little bit better feedback but most of the assignments I did were correct."

### Sentiment Analysis

The sentiment analysis provided a quantitative measure of student engagement across the survey’s open-ended responses. Using the NRC sentiment lexicon (implemented via the syuzhet package), we quantified positive, negative, and neutral sentiments for each question.

The analysis revealed:

- **Predominantly Positive Sentiment**: Most responses expressed positive sentiments, especially regarding structured content delivery, effective use of R, and clear statistical explanations.

- **Mixed Feedback on Clarity and Pace**: While many responses were positive overall, some negative sentiments were associated with challenges such as rapid topic progression and difficulties with R syntax.

- **Critical Comments**: Negative terms like “confusing,” “fast,” and “examples” frequently appeared in responses about areas needing clearer instructional support.

The bar chart above (generated in the sentiment analysis code chunk) illustrates the distribution of sentiment (positive, negative, neutral) for each of the 8 survey questions.

```{r sentiment-plot, echo=FALSE, fig.cap="Sentiment Distribution by Survey Question"}
# Reshape the data to long format for plotting
sentiment_long <- sentiment_results %>%
  pivot_longer(cols = c("Positive", "Negative", "Neutral"), 
               names_to = "Sentiment", 
               values_to = "Count")

# Set correct order for x-axis
sentiment_long$Question <- factor(sentiment_long$Question,
                                  levels = question_names)

# Plot
ggplot(sentiment_long, aes(x = Question, y = Count, fill = Sentiment)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("Positive" = "darkgreen", 
                               "Negative" = "red", "Neutral" = "gray")) +
  theme_minimal() +
  labs(title = "Sentiment Distribution by Survey Question",
       x = "Survey Question",
       y = "Sentiment Count")
```

# Conclusion

The STA107 post-course survey provides valuable insights into how students perceive the course structure and its technical components. Findings highlight both effective practices and opportunities for enhancement. Continued evaluation and iteration will ensure the course remains responsive to student needs and pedagogical best practices.

From a qualitative perspective, student feedback revealed strong appreciation for several key elements of the course especially the hands-on R activities, the integration of real-world statistical examples, and the clarity of normal model explanations. These components were described as engaging and instructive, affirming their value in helping students apply theoretical knowledge in practical settings.

However, when students were asked about areas of improvement, many raised concerns regarding the speed at which complex topics were introduced, as well as the difficulty of interpreting R-based assignments without more structured guidance. Themes of pacing, clarity, and the need for more scaffolded instruction were particularly prominent. Addressing these concerns by slowing down the delivery of technical material, providing more detailed examples, and incorporating mid-semester feedback opportunities could significantly enhance the learning experience in future offerings of STA107.

# References

- Jockers, M. (2017). *syuzhet: Extracts Sentiment and Sentiment-Derived Plot Arcs from Text* (R package v1.0.4). [CRAN Documentation](https://cran.r-project.org/web/packages/syuzhet/index.html)

- Jockers, M. (2020). *Introduction to the Syuzhet Package*. [CRAN Vignette](https://cran.r-project.org/web/packages/syuzhet/vignettes/syuzhet-vignette.html)

- Kim, H. (2022). Sentiment Analysis: Limits and Progress of the Syuzhet Package and Its Lexicons. *Digital Humanities Quarterly, 16*(2). [Article](http://www.digitalhumanities.org/dhq/vol/16/2/000601/000601.html)